---
title: "Chr21 Population Structure"
author: "Harrison Ostridge"
date: "16/06/2021"
output: html_document
---

The purpose of this script is to plot PCAs from the f7 filtered chr21 data to investigate population structure in chr21 samples. I want to verify that they agree with previous studies and the exomes.

#### Setup
```{r setup}
rm(list=ls())
knitr::opts_knit$set(root.dir = '~/OneDrive - University College London/Projects/Ostridge_PanAf/population_structure/') 
```

#### Library
```{r library, message=FALSE, warning=FALSE}
library(dplyr)
library(data.table)
library(ggplot2)
library(ggrepel)
library(readxl)
library(plyr)
library(tidyr)
library(stringr)
library(igraph)
library(RColorBrewer)
library(grid) 
library(gridExtra)
library(fields)
library(maps)
library(mapdata)
library(mapplots)
library(RColorBrewer)
library(MCMCpack)

filtering_dir="../sample_filtering/chr21/output/"
pcangsd_output_dir="../pcangsd/chr21/output/"
```

#### Read in data

```{r}
# Read in metadata
f7_chr21=read.csv(paste0(filtering_dir, "/f7/f7.chr21.metadata.csv"), check.names = FALSE)
f7_exome=read.csv(paste0(filtering_dir, "/f7/f7.metadata.csv"), check.names = FALSE)
## Ensure it is in the correct order
f7_chr21=f7_chr21[order(f7_chr21$Sample),]
f7_exome=f7_exome[order(f7_exome$Sample),]
```

## Population Structure

### PCAngsd

```{r}
plot.PCA=function(subsps, meta.data, input.dir, output.file=NULL, label=NULL, return.df=FALSE){
  # Create list to store result data frames in
  results=list()
  # BAM file lists are always given in alphabetical order (with respect to sample name) so it is important to ensure the samples are in this order
  meta.data=meta.data[order(meta.data$Sample),]
  # Add label column, if no column is provided a blank label column is used
  meta.data$label=''
  if(!is.null(label)){meta.data$label=meta.data[[label]]}
  # Loop over each subspecies provided
  for(subsp in subsps){
    title=""
    meta.data.subsp=meta.data
    # Subspecies options
    if(subsp=="c"){title="Central"}
    if(subsp=="e"){title="Eastern"}
    if(subsp=="n"){title="Nigeria-Cameroon"}
    if(subsp=="w"){title="Western"}
    if(subsp=="all"){title="All Samples"
      # Group defines how to draw polygons
      group="Subspecies"
      input=paste0(input.dir,"chr21.f7.", subsp,".cov")}
    if(subsp != "all"){
      # Group defines how to draw polygons
      group="Site"
      input=paste0(input.dir, subsp,"/chr21.f7.", subsp,".cov")
      if(title!=""){meta.data.subsp=meta.data[meta.data$Subspecies==title,]}}
    # Read in PCAngsd output
    cov=read.table(input)
    pcs=eigen(cov)
    # Select columns corresponding to PC1 and PC2
    pc1.pc2=as.data.frame(pcs$vectors[,1:2])
    colnames(pc1.pc2)=c("f7.PC1", "f7.PC2")
    # Add PC1 and PC2 columns to metadata
    ## The order of samples in PCAngsd output should be alphabetical as the BAM file lists were written in this order
    pc1.pc2=cbind(meta.data.subsp, pc1.pc2)
    # Percentage variance explained by PCs
    PC1.percent=pcs$values[1]/sum(pcs$values)*100
    PC2.percent=pcs$values[2]/sum(pcs$values)*100
    # Plot
    PC1.lab=paste("PC1 (", round(PC1.percent,digits = 2),"%)")
    PC2.lab=paste("PC2 (", round(PC2.percent,digits = 2),"%)")
    find_hull=function(pc1.pc2) pc1.pc2[chull(pc1.pc2[,'f7.PC1'], pc1.pc2[,'f7.PC2']), ]
    hulls=ddply(pc1.pc2, group, find_hull)
    # Plot all samples
    if(subsp=="all"){
      plot=ggplot(pc1.pc2, aes(x=f7.PC1, y=f7.PC2, fill=Subspecies, colour=Subspecies, label = label)) +
        theme_minimal()+ 
        geom_point(aes(shape=Subspecies), size = 2) + 
        geom_text_repel(size = 3, min.segment.length = unit(0.1, "lines"), show.legend = FALSE) +
        geom_polygon(data = hulls, alpha = 0.5) + 
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key.size = unit(0.4, "cm"), plot.title = element_text(hjust = 0.5)) +
        guides(fill=guide_legend(ncol=1)) + 
        labs(x=paste(PC1.lab), y=paste(PC2.lab)) + 
        ggtitle(title) +
        scale_fill_manual(values = c('green3', 'darkorange', 'red', 'blue')) +
        scale_color_manual(values = c('green3', 'darkorange', 'red', 'blue')) +
        scale_shape_manual(values=0:4)
    }
    # Plot individual subspecies
    else{
      plot=ggplot(pc1.pc2, aes(x=f7.PC1, y=f7.PC2, fill=Site, colour=Site, label = label)) +
        theme_minimal()+ 
        geom_point(aes(shape=Site), size = 2) + 
        geom_text_repel(size = 3, min.segment.length = unit(0.1, "lines"), show.legend = FALSE) +
        geom_polygon(data = hulls, alpha = 0.5) + 
        theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key.size = unit(0.4, "cm"), plot.title = element_text(hjust = 0.5)) +
        guides(fill=guide_legend(ncol=1), shape=guide_legend(ncol=1)) + 
        labs(x=paste(PC1.lab), y=paste(PC2.lab)) + 
        ggtitle(title) +
        scale_shape_manual(values=rep(0:6, 10))
    }
    # If an output file is provided, write to it
    if(!is.null(output.file)){ggsave(paste0(output.file), plot=plot)}
    # Display plot
    print(plot)
    # Save metadata with PC1 and PC2 columns added for the subspecies in results list
    results[[subsp]]=pc1.pc2
  }
  # If return.df is set to TRUE, return the results list
  if(return.df==T){return(results)}
}
```

```{r message=FALSE, warning=FALSE, out.width="50%"}
plot.PCA("all", 
         f7_exome,
         paste0(pcangsd_output_dir, "/chr21.f7.all_minInd.15_doMajorMinor.1_HWE.p.1e-3_snp.p.1e-6_minMAF.0.01/"))
```


```{r  message=FALSE, warning=FALSE, out.width="50%"}
plot.PCA(c("c", "e", "n", "w"), 
         f7_exome,
         paste0(pcangsd_output_dir, "/chr21.f7.subsp_minInd.15_doMajorMinor.1_HWE.p.1e-3_snp.p.1e-6_minMAF.0.01/"))
```

### Easetrn PCA looks rather different in chr21 compared to exomes - somthing to bear in mind 

- This chr21 one looks more like the one in Claudia's paper than the exome one unsurprisingly 

## Procrustes transformation

```{r out.width="50%"}
plot.PCA.map=function(subsps, meta.data, input.dir, output.file=NULL, label=NULL, return.df=FALSE){
  # Create list to store result data frames in
  results=list()
  # BAM file lists are always given in alphabetical order (with respect to sample name) so it is important to ensure the samples are in this order
  meta.data=meta.data[order(meta.data$Sample),]
  # Loop over each subspecies provided
  for(subsp in subsps){
    title=""
    meta.data.subsp=meta.data
    # Subspecies options
    if(subsp=="c"){title="Central"}
    if(subsp=="e"){title="Eastern"}
    if(subsp=="n"){title="Nigeria-Cameroon"}
    if(subsp=="w"){title="Western"}
    if(subsp=="all"){title="All Samples"
      # Group defines how to draw polygons
      group="Subspecies"
      input=paste0(input.dir,"chr21.f7.", subsp,".cov")}
    if(subsp != "all"){
      # Group defines how to draw polygons
      group="Community"
      input=paste0(input.dir, subsp,"/chr21.f7.", subsp,".cov")
      if(title!=""){meta.data.subsp=meta.data[meta.data$Subspecies==title,]}}
    # Read in PCAngsd output
    cov=read.table(input)
    pcs=eigen(cov)
    # Procrustes transformation
    pro=procrustes(pcs$vectors[,1:2], as.matrix(meta.data.subsp[, c("Longitude", "Latitude")]),translation=TRUE,dilation=TRUE)$X.new
    colnames(pro)=c("trans.long", "trans.lat")
    pro.meta.data=cbind(meta.data.subsp, pro)
    # Hulls
    find_hull=function(pro.meta.data) pro.meta.data[chull(pro.meta.data[,'trans.long'], pro.meta.data[,'trans.lat']), ]
    hulls=ddply(pro.meta.data, group, find_hull)
    # Limits
    ## x
    xmin=min(min(pro.meta.data$trans.long), min(pro.meta.data$Longitude))
    xmax=max(max(pro.meta.data$trans.long), max(pro.meta.data$Longitude))
    xmin=xmin-(xmax-xmin)*0.2
    xmax=xmax+(xmax-xmin)*0.2
    ## x
    ymin=min(min(pro.meta.data$trans.lat), min(pro.meta.data$Latitude))
    ymax=max(max(pro.meta.data$trans.lat), max(pro.meta.data$Latitude))
    ymin=ymin-(ymax-ymin)*0.1
    ymax=ymax+(ymax-ymin)*0.1
    # Line connecting polygon centroids to geographic locations
    centroids=aggregate(as.matrix(pro.meta.data[,c("trans.long", "trans.lat")]) ~ Site, data=pro.meta.data, FUN=mean)
    colnames(centroids)=c("Site","cent.long", "cent.lat")
    lines=merge(centroids, unique(pro.meta.data[,c("Site","Longitude", "Latitude")]))
    # Plot
    world_map=map_data("world")
    plot=ggplot(data=pro.meta.data, aes(x=trans.long, y=trans.lat, fill=Site, colour=Site)) +
      theme_minimal() +
      geom_polygon(data=world_map, aes(x = long, y = lat, group = group), fill="grey93", colour = "darkgrey") +
      coord_sf(xlim=c(xmin, xmax), ylim=c(ymin, ymax)) +
      geom_point(aes(shape=Site), size = 1) + 
      geom_point(aes(x=Longitude, y=Latitude, shape=Site), size = 4) + 
      geom_polygon(data = hulls, alpha = 0.5) + 
      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key.size = unit(0.4, "cm"), plot.title = element_text(hjust = 0.5)) +
      guides(fill=guide_legend(ncol=1), shape=guide_legend(ncol=1)) + 
      geom_segment(data = lines, aes(x = Longitude, y = Latitude, xend = cent.long, yend = cent.lat, colour = Site))+
      ggtitle(title) +
      scale_shape_manual(values=rep(0:6, 10)) +
      geom_text_repel(data=lines, aes(x=Longitude, y=Latitude, label=Site), 
                        size = 3, segment.color="black", segment.linetype=1, segment.size=0.2, show.legend = FALSE, fontface = 'bold', colour="black") +
      xlab("") + ylab("")
    print(plot)
  }
}

plot.PCA.map(c("c", "e", "n", "w"), 
         f7_exome,
         paste0(pcangsd_output_dir, "/chr21.f7.subsp_minInd.15_doMajorMinor.1_HWE.p.1e-3_snp.p.1e-6_minMAF.0.01/"))
```
